---
title: "assignment-2"
author: "Gabriele Chignoli"
date: "10/9/2017"
output: html_document
---

## Exercise 1: Writing R functions

All functions written for this assignment will be provided from the source file:
```{r}
source("functions.R")
```

You will find there all commentaries on functions, what they do and why.

### Question 1a: filling in the blanks

The codes below provide the results of the first filling.
In order to verify the presence of numeric values in a certain column of the dataset and then to sum them all, an ```If``` condition was added to the function ```sum_column```:

```{r}
print(sum_column(iris, "Sepal.Length"))
print(sum_column(iris, "Species"))
print(sum_column(warpbreaks, "breaks"))
```

### Question 1b: a function with one argument and one return value

As a result of cheating on first filling by adding the built-in **R** ```sum()```, here we use a basic code to recreate a function which can do the sum for us. We test it using the code chunk below:

```{r}
print(my_sum(iris$Sepal.Length))
print(my_sum(iris$Species))
print(my_sum(warpbreaks$breaks))
```

We lately add this function to the first one.

### Question 1c: a function with two arguments and one return value

This time we have a new function to make, it uses our sum function too and it divides the sum for a number. As done earlier here is a chunk in order to verify the result:

```{r}
print(sum_divided_by(iris$Sepal.Length, 12))
print(sum_divided_by(iris$Species, 22))
print(sum_divided_by(iris$Sepal.Length, "Not numeric"))
print(sum_divided_by(warpbreaks$breaks, -12))
```

### Question 1d: one more function

Last function we create for this exercise calculates the mean of a vector not cheating using the built-in ```mean()```. Chunk below show the result:

```{r}
print(my_mean(iris$Sepal.Length))
print(my_mean(iris$Species))
print(my_mean(warpbreaks$breaks))
```

## Exercise 2: Working with ggplot

Here we'll work with the ```ggplot``` package and as we did earlier all functions and code will be provided in

```{r}
source("functions.R")
```

### Question 2a: creating violin plots

We'll test this first question with the following chunk:

```{r}
print(grouped_violin_plot(iris, "Sepal.Length", "Species"))
```

### Question 2b: modifying plots

This time we will just modify the last plot in order to add a main title and to change plot colours, thanks to the chunk you see here:

```{r}
p <- grouped_violin_plot(iris, "Sepal.Length", "Species")
p <- p + ggplot2::scale_fill_brewer(palette="Accent")
p <- p + ggplot2::labs(title="Iris data")
print(p)
```

## Exercise 3: Permutation tests

As we did in Exercise 1, here we will add some new functions to our `functions.R` file and every new line will come with commentaries to explain what we wanted to obtain or what we do.

### Question 3a: writing a more generic function for taking a test statistic

Our first function calculate the difference between medians of a selected group. Median correspond to the middle value of a sequence, in our case of our column, but what happens when the sequence is unpair ? We take the two middle values and calculate the halfway between them, that's what happened to us, where our columns are length 50.  
In **R** a built-in function provide to estimate the median value, but we had to do some adjustments before calling it because our column presented a String list of numeric values.  
Once we did everything necessary to make possible our computing, we obtain :  

* Versicolor median = 2.8  
* Virginica median = 3.0  

```{r}
difference_in_medians(iris, "Sepal.Width", "Species", "versicolor", "virginica")
difference_in_medians(iris, "Sepal.Width", "Species", "virginica", "virginica")
```

###Question 3b: writing a more generic function of the randomization (i.e., permutation) function

This time we work on columns too but instead of using values in columns to obtain a result, we take all the column dataset and randomly replace them.

```{r}
iris$Sepal.Width[1:10]
if(!exists(".Random.seed")) set.seed(NULL)
previous_seed <- .Random.seed
set.seed(1)
randomize(iris, "Sepal.Width")$Sepal.Width[1:10]
randomize(iris, "Species")$Species[1:10]
randomize(iris, "Species")$Sepal.Width[1:10]
set.seed(previous_seed)
```