---
title: "assignment-2"
author: "Gabriele Chignoli"
date: "10/9/2017"
output: html_document
---

## Exercise 1: Writing R functions

All functions written for this assignment will be provided from the source file:
```{r}
source("functions.R")
```

You will find there all commentaries on functions, what they do and why.

### Question 1a: filling in the blanks

The codes below provide the results of the first filling.
In order to verify the presence of numeric values in a certain column of the dataset and then to sum them all, an ```If``` condition was added to the function ```sum_column```:

```{r}
print(sum_column(iris, "Sepal.Length"))
print(sum_column(iris, "Species"))
print(sum_column(warpbreaks, "breaks"))
```

### Question 1b: a function with one argument and one return value

As a result of cheating on first filling by adding the built-in **R** ```sum()```, here we use a basic code to recreate a function which can do the sum for us. We test it using the code chunk below:

```{r}
print(my_sum(iris$Sepal.Length))
print(my_sum(iris$Species))
print(my_sum(warpbreaks$breaks))
```

We lately add this function to the first one.

### Question 1c: a function with two arguments and one return value

This time we have a new function to make, it uses our sum function too and it divides the sum for a number. As done earlier here is a chunk in order to verify the result:

```{r}
print(sum_divided_by(iris$Sepal.Length, 12))
print(sum_divided_by(iris$Species, 22))
print(sum_divided_by(iris$Sepal.Length, "Not numeric"))
print(sum_divided_by(warpbreaks$breaks, -12))
```

### Question 1d: one more function


```{r}
print(my_mean(iris$Sepal.Length))
print(my_mean(iris$Species))
print(my_mean(warpbreaks$breaks))
```